Lucky Thirteen - A Cryptographic Analyses

### 2013-07-10 paper.tex
Spellchecked Section II

\begin{abstract}

@@

\section{Brief overview of Cryptography}
\label{sec:crypto}
In TLS, different cryptographic primitives are used to provide confidentiality, authentication, non-repudiation and integrity while communicating over the internet. Since TLS uses a MAC-Encode-Encrypt Scheme, the different subsections will describe these processes each. These subsections are written with TLS in mind, and cover specifics for the implementation of TLS rather than a full description of the specific cryptographic primitive. The first subsection denotes the cryptographic format that is used in this paper.

\subsection{Cryptographic format}
\label{sec:crypto:format}
The cryptographic format used in this paper is presented in Table~\ref{sec:crypto:format:table}.
\begin{table}[h]
    \begin{tabular}{ll}
    Format & Meaning \\
    A $||$ B    & Concatenation of A and B  \\
    A $\oplus$ B  & The bitwise XOR of A and B     \\
    H(A)    & Hash-function applied to A    \\
    $E_k$(A)    & Encryption with key K applied to A    \\
    \end{tabular}
    \caption{Cryptographic format used in this paper}
    \label{sec:crypto:format:table}
\end{table}

\subsection{MAC - Message Authentication Code}
\label{sec:crypto:hmac}
To ensure message integrity, RFC5246~\cite{ietf2008transport} prescribes the use of a keyed Message Authentication Code (MAC). The MAC cipher suites described in the RFC are based on hashing functions, HMAC, and are described in~\cite{krawczyk1997rfc}. Though, the RFC does not prohibit the use of other cipher suites, the implementations found by~\citeauthor{alfardan2013lucky} in~\cite{alfardan2013lucky} all restrict the use to HMAC-MD5 ($16$ bytes of MAC), HMAC-SHA-1 ($20$ bytes of MAC) or HMAC-SHA-256 ($32$ bytes of MAC).

To compute the HMAC for a message M and key K, the specified HMAC algorithms use the following formula. 
\[ MAC = H((K \oplus opad)||H(K \oplus ipad)||M)). \]
With opad, ipad to be specific $64$-byte values and key K zero-padded to $64$ bytes. "Each hash function uses a \textit{Merkle-Damg\aa rd strengthening}. Here, an 8-byte length field followed by padding of a specified byte format are appended to the message M to be hashed. The padding is at least $1$ byte in length and aligns the data on a $64$-byte boundary. The relevant hash functions also have an iterated structure, processing messages in chunks of $64$ bytes ($512$ bits) using a compression function, with the output of each compression step being chained into the next step. The compression function in turn involves a complex round structure, with many basic arithmetic operations on data being involved in each round."~\cite{alfardan2013lucky}. This means that, depending on the size of message M, the number of runs of the hash function H can be computed. Because of the Merkle-Damg\aa rd strengthening, a message M is appended by the $8$-byte length field, and padded to be a multiple of $64$ bytes. This means that the first of the Merkle-Damg\aa rd rounds consists of $55$ bytes of message M, followed by $64$ bytes for every following $64$ bytes of message M. The inner hash function depends on the message length of M $+ 1$ constant hash function round, the outer hash function is a constant of $2$. This means that the following formula represents the number of hash function calls for L is the message length of M in bytes:
\[ \lceil \frac{L - 55}{64} \rceil + 4 \]

\subsection{Encoding - Padding}
\label{sec:crypto:padding}
Not specifically a cryptographic primitive, though, padding forms a core point within the different attacks explained within this paper. TLS describes the padding to be at least $1$ byte in length, with the content being the number of padding bytes $- 1$. This forces padding such as:
\["0x00", "0x01||0x01", "0x02||0x02||0x02", etc. \]

\subsection{Encryption - CBC encryption}
\label{sec:crypto:encryption}
The attack described by~\citeauthor{alfardan2013lucky}~\cite{alfardan2013lucky} make use of one the most prominent implementation of RFC~\cite{ietf2008transport} (TLS 1.1 and 1.2), namely the use of a Block Cipher in CBC-mode with a randomized IV. In practice, AES is used as the Block Cipher. The plaintext is cut into blocks of the input size of the Block Cipher (AES has a 16 bytes block size). Ciphertext blocks are computed as:
\[ C_j = E_K(P_j \oplus C_{j-1}) \]
With $P_i$ the blocks of plaintext (and extras, described in Section~\ref{sec:TLS}), $C_0$ the IV and $K$ the key for Block Cipher $E$. Ciphertext C is the concatenation of all $C_i$'s.

>>>>

\begin{abstract}
TODO.

@@

\section{Brief overview of Cryptography}
\label{sec:crypto}
In TLS, different cryptographic primitives are used to provide confidentiality, authentication, non-repudiation and integrity while communicating over the internet. Since TLS uses a MAC-Encode-Encrypt Scheme, the different subsections of this section will describe these processes. These subsections are written with TLS in mind, and cover specifics for the implementation of TLS rather than a full description of the specific cryptographic primitive. The first subsection denotes the cryptographic format that is used in this paper.

\subsection{Cryptographic format}
\label{sec:crypto:format}
The cryptographic format used in this paper is presented in Table~\ref{sec:crypto:format:table}.
\begin{table}[h]
    \begin{tabular}{l|l}
    Format & Meaning \\ \hline
    A $||$ B    & Concatenation of A and B  \\
    A $\oplus$ B  & The bitwise XOR of A and B     \\
    H(A)    & Hash-function applied to A    \\
    $E_k$(A)    & Encryption with key K applied to A    \\
    \end{tabular}
    \caption{Cryptographic format used in this paper}
    \label{sec:crypto:format:table}
\end{table}

\subsection{MAC - Message Authentication Code}
\label{sec:crypto:hmac}
To ensure message integrity, RFC5246~\cite{ietf2008transport} prescribes the use of a keyed Message Authentication Code (MAC). The MAC cipher suites described in the RFC are based on hashing functions, HMACs, and are described in~\cite{krawczyk1997rfc}. Though, the RFC does not prohibit the use of other cipher suites, the implementations found by~\citeauthor{alfardan2013lucky} in~\cite{alfardan2013lucky} all restrict the use to HMAC-MD5 ($16$ bytes of MAC), HMAC-SHA-1 ($20$ bytes of MAC) or HMAC-SHA-256 ($32$ bytes of MAC).

To compute the HMAC for a message M and key K, the specified HMAC algorithms use the following formula. 
\[ HMAC(K,M) = H((K \oplus opad)||H(K \oplus ipad)||M)). \]
With opad, ipad to be specific $64$-byte values and key K zero-padded to $64$ bytes. "Each hash function uses a \textit{Merkle-Damg\aa rd strengthening}. Here, an 8-byte length field followed by padding of a specified byte format are appended to the message M to be hashed. The padding is at least $1$ byte in length and aligns the data on a $64$-byte boundary. The relevant hash functions also have an iterated structure, processing messages in chunks of $64$ bytes ($512$ bits) using a compression function, with the output of each compression step being chained into the next step. The compression function in turn involves a complex round structure, with many basic arithmetic operations on data being involved in each round."~\cite{alfardan2013lucky}. This means that, depending on the size of message M, the number of runs of the hash function H can be computed. Because of the Merkle-Damg\aa rd strengthening, a message M is appended by the $8$-byte length field, and padded to be a multiple of $64$ bytes. This means that the first of the Merkle-Damg\aa rd rounds consists of $55$ bytes of message M, followed by $64$ bytes for every following $64$ bytes of message M. The inner hash function depends on the $message length of M + 1$ constant hash function round, the outer hash function is a constant of $2$. This means that the following formula represents the number of hash function calls for L is the message length of M in bytes:
\[ \lceil \frac{L - 55}{64} \rceil + 4 \]

\subsection{Encoding - Padding}
\label{sec:crypto:padding}
Not specifically a cryptographic primitive, though, padding forms a core point in the different attacks explained within this paper. TLS describes the padding to be at least $1$ byte in length, with the content being the $number of padding bytes - 1$. This forces padding such as:
\["0x00", "0x01||0x01", "0x02||0x02||0x02", etc. \]

\subsection{Encryption - CBC encryption}
\label{sec:crypto:encryption}
The attack described by~\citeauthor{alfardan2013lucky}~\cite{alfardan2013lucky} makes use of one the most prominent implementations of the RFC~\cite{ietf2008transport}, namely the use of a Block Cipher in CBC-mode with a randomized IV. In practice, AES is used as the Block Cipher.

The plaintext is cut into blocks of the input size of the Block Cipher (AES has a $16$ bytes block size). Ciphertext blocks are computed as:
\[ 
\begin{split}
C_i &= E_K(P_i \oplus C_{i-1}) \\
C_0 &= IV 
\end{split}
\]
With $P_i$ the blocks of plaintext (plus extras, as described in Section~\ref{sec:TLS}) and $K$ the key for Block Cipher $E$. Ciphertext $C$ is the concatenation of all $C_i$'s.
